A practical guide to Digital Micro-mirror Devices (DMDs) for wavefront shaping

### 2023-11-21 paper/main.tex
Updates from Overleaf

\documentclass[12pt]{iopart}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

@@

Digital Micromirror Devices (DMDs) have emerged as a technology bridging the gap
between these two types of systems;
they offer a large number of pixels (similar to LC-SLMs) and fast modulation speeds (typically up to several tens of kHz).
Their high speed capabilities made them attractive for real-time applications,
in particular for high-resolution imaging microscopy
requiring fast scanning or illumination shaping~\cite{cha2000nontranslational,Zhuang2020},
biolithography~\cite{yoon2018emerging},
and optical tweezers~\cite{gauthier2016direct}.
However, DMDs are restricted to hardware binary amplitude modulation and are not optimized for coherent light applications.
Utilizing DMDs for coherent control of light in complex media is therefore non-trivial
and necessitates specific adaptations for efficient use.\\

To comprehend both the capabilities and limitations
of DMD technology for coherent wavefront shaping,
it is crucial to understand the device's operating principles
and its original design intentions.
Investigated and developed by Texas Instruments since the 1980s,
DMDs gained prominence in the 1990s for video projection applications since the 1990s
under the commercial name of Digital Light Processing (DLP)~\cite{hornbeck1997digital,Dudley2003emerging}.
The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
DMDs operate by toggling the state of small mirrors between two distinct angles, denoted as $\pm \theta_\text{DMD}$.
The device is originally engineered for amplitude modulation in video projection applications.
In this configuration, one mirror angle directs light into the projection lens,
while the alternate angle results in the light path being blocked (see Fig.~\ref{fig:combined_pixel}).
Given that projectors utilize incoherent light and that the DMD plane is optically conjugated with the projection screen,
aberrations within the DMD plane are generally not problematic.
Similarly, phase fluctuations induced by temperature variations,
as well as minor vibrations from the cooling hardware, are inconsequential in this context.
The DMD is designed to produce binary on/off modulation,
which is then leveraged to generate grayscale images via pulse-width modulation.
Color modulation is accomplished through the use of a color wheel in conjunction with a bright white light source.\\

@@

The 2D array can be seen as 2 orthogonal 1D gratings in the $x$ and $y$ directions,
both having the same properties.
A more comprehensive depiction of the 2D system can be found in~\cite{Scholes2019structured}.
In order to utilize Eq.(\ref{eq:blazed_eq}),
one first needs projecting
the different angles of the problem onto the incident planes
of the two 1D gratings.
This yields
$\alpha = \text{arctan}\left(\text{tan}(\alpha')/\sqrt{2}\right)$,
$\beta = \text{arctan}\left(\text{tan}(\beta')/\sqrt{2}\right)$,
and $\theta_B = \text{arctan}\left(\text{tan}(\theta_\text{DMD})/\sqrt{2}\right)$,
where $\theta_\text{DMD}$ represents the true rotation angle of the mirrors relative to the diagonal of the pixels,
and $\alpha'$ and $\beta'$ denote the incident and outgoing angles in the horizontal plane.
We can quantify how close we are to the ideal case,
i.e., when satifying the blazed equation outlined by Eq.(\ref{eq:blazed_eq}),
by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:

@@

with $\lfloor . \rfloor$ representing the integer part,
and $\mod{2}$ the modulo 2 operation.
$\mu$ is maximal and equals $1$ when the blazing equation is satisfield,
i.e. when one order of diffraction contains most of the energy,
and minimal when we are in the worst case scenario,
i.e. when four orders of diffraction have a significant and equal intensity.\\

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/mu_76.pdf}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/mu_108.pdf}
  \end{subfigure}
  \caption{
    \textbf{Blazed number and far field diffraction patterns, extreme cases.}
    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha'$  (top)
    for a pixel pitch of $d=7.6$\textmu m (left)
    and $d=10.8$\textmu m (right).
    Corrseponding far field diffraction pattern (bottom)
    for two incident angles $\alpha' = 0^\circ$ and $\alpha' = 25^\circ$.
    The yellow cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
  }
  \label{fig:mu76}
\end{figure}

@@

\begin{tldr}
  \textbf{TL;DR:}\\
  DMDs act as a blazed gratings.
  For a given operation wavelength,
  we need to find the right pixel pitch
  to have a good modulation quality
  and diffraction efficiency.
  It can be done by estimating the \textit{blazed number} $\mu$ introduced in eq.~(\ref{eq:blazed_number})
  or directly using our custom online tool~\cite{popoffDMDDiffractionTool}.
\end{tldr}

@@

While only capable of providing hardware binary amplitude modulation,
DMDs serve as an potent tool for wavefront shaping and sensing.
These applications  critically require characterization and correction of aberrations
caused by the non-flatness of the DMD surface.
For LC-SLMs, the manufacturer typically characterizes the surface's inhomogeneities within the plane of the modulator and
provides a spatial phase profile of the introduced aberrations based on the operational wavelength.
It is noteworthy that when utilized for intensity modulation on a plane conjugated to the DMD plane,
as it is the done in digital projectors,
the system becomes insensitive to the aberrations caused by the DMD surface.
Consequently, these effects are commonly overlooked and rarely documented in the information provided by the manufacturers.
They are nonetheless non-negligible, with deviations from a flat surface
that typically spans across several wavelengths~\cite{Brown2021multicolor}.\\

@@

In the literature, various methods have been proposed to characterize
the phase pattern of the DMD aberrations in the plane of the modulator.
Typically, this involves the use of a model for the aberrations,
tweaking the parameters to align with the measurements~\cite{Matthes2019Optical,Scholes2019structured, Brown2021multicolor}.
An alternative approach entails direct measurement of the distorted wavefront,
either employing a wavefront sensor such as a Shack-Hartmann~\cite{Lee2023compensation},
or via interferometry.
While such methods yield accurate results,
they necessitate adaptation to the particular
setup conditions,
and frequently require supplementary optical components,
meticulous alignment,
and custom software.
Furthermore, they also demand
a meticulously calibrated and stable setup.
In this section, we introduce a straightforward method to characterize the aberrations using a lens and a camera.
This technique can be employed for any system that offers phase modulation,
such as LC-SLMs or deformable mirrors. \\

@@

We assume that the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}.
This implies that the modulator can be divided into $N$ sections,
which we designate as {\em macropixels},
where the phase can be controlled independently.
We use a lens and a camera in its Fourier plane,
illuminating the modulator with a collimated beam
that extends over the entire area of the modulator intended for use.
In scenarios where there are minimal or no aberrations,
the intensity pattern observed would mimic the PSF of the lens,
such as an Airy disk depicted in Fig.~\ref{fig:dmd_aberr_setup}.c.
However, in practice, we encounter a substantially distorted pattern,
like the one represented in Fig.~\ref{fig:dmd_aberr_setup}.d.
A more detailed depiction of the setup for aberration characterization,
in the context of a wavefront shaping application in complex media is presented in Fig.~\ref{fig:dmd_aberr_setup2}.a.
In the case where the medium is a multimode fiber,
one could leverage the reflection from the input surface
to directly visualize the intensity pattern of the input plane,
as shown in Fig.~\ref{fig:dmd_aberr_setup2}.b.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/aberr_setup.pdf}
  \caption{
    \textbf{Detailed aberration characterization setup.}
    (a) A laser beam is expanded and collimated using a telescope (lenses L and L').
    The DMD is represented in transmission for simplicity.
    Light reflected from the DMD is focused by a lens L$_\text{1}$
    onto the complex medium to study, namely a scattering medium (SC)
    or a multimode fiber (MMF).
    A beamsplitter is used to image the far field of the DMD onto a camera
    using a lens (L$_\text{2}$).
    The intensity pattern, up to a homotetic transformation,
    represent the input excitation on the mediun
    which also correspond to the PSF of the illumination setup.
    (b) In the case of a multimode fiber,
    one can take advantage of the reflection of the input facet
    to directly image the intensity pattern of the input plane.
  }
  \label{fig:dmd_aberr_setup2}
\end{figure}

@@

As an illustration, we conduct an optimization procedure using 11 Zernike polynomials.
We use a V-9501 Vialux DMD with a DLP9500 TI chip
of resolution 1920 by 1200 pixels and a pixel pitch of 10.8 \textmu m.
The optimization is performed on a disk of radius $340$ pixels,
The illumination is done using an expanded laser beam at 633 nm,
corresponding to the aperture of our optical setup.
We exclude the first three Zernike polynomials in the optimization process,
starting from the radial degree $2$.
Indeed, the initial one, known as the piston, has no influence on the PSF quality.
The subsequent ones, the tip and the tilt, cause the PSF to shift.
Our procedure relies on optimizing the maximum of the PSF, wherever that is,
rendering us indifferent to these two parameters.
After optimization,
it is possible to generate the correction pattern using a selected number of Zernike polynomials
in order to investigate their impact on the PSF quality.
We see here that using about 10 Zernike polynomials is sufficient to obtain a Strehl ratio $>0.99$.
Fig.~\ref{fig:zernike} demonstrates the Strehl ratio and the intensity profiles of the PSF
for different counts of the utilized Zernike polynomials.
It is important to note that we do not use the full surface of the DMD.
Using a larger area may lead to stronger deformations of the PSF,
requiring a larger number of Zernike polynomials to be corrected accurately.
The experimental data, in addition to the Python code used to generate the figures, can be accessed in the dedicated repository~\cite{github}.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/Zernike_2.pdf}
  \caption{
    \textbf{Effect of Zernike polynomials on the optimizaton of the PSF.}
    (a) Strehl ratio of the point spread function as a function of the number of Zernike polynomials used
    for the compensation of the aberrations.
    (b) Intensity profile of the point spread function for different number of Zernike polynomials
    in linear and logarithmic scale.
    To generate this data, we take the final solution of the optimization experimental procedure,
    i.e. the optimal coefficients $a'_n$,
    and generate the corrected PSF with an increasing number of Zernike polynomials.
  }
  \label{fig:zernike}
\end{figure}

@@

Unlike the original purpose of the DMD, i.e. ampltiude modulation for video projectors,
typical scientific applications require a high stability of the generated wavefront.
This is particularly true for applications in complex media,
such as strongly scattering media or multimode fiber,
where a small change in the phase front can lead to a large change in the output intensity profile.
While LC-SLMs design has been improved and adapted to scientific applicaitons
over the last decades,
DMD are still a relatively new tools for wavefront shaping and sensing
and are prone to instabilities that need to be adressed by the user.
In this section we present the effect of mechanical and thermal instabilities
and how to limit their impact on the wavefront quality with simple solutions.\\

@@

In the supplementary materials~\cite{SI}, we present an animation illustrating the dynamic pattern.
In off-axis holography,
the local transverse displacement of the fringes
is directly proportional to the phase,
with a displacement equivalent to the period of the fringes corresponding to $2\pi$~\cite{Cuche2000spatial}.
This permits us to estimate the fluctuation in phase over time at a given postition of the output plane.
As illustrated in Fig.~\ref{fig:phase_vibrations} (depicted by the red curve),
the phase varies rapidly over time,
a fluctuation attributed to the rapid mechanical vibrations transmitted by the board.\\

@@

An simple yet effective solution consists in dampening the vibrations
at the flat cable's level by clamping it with a soft material,
as depicted in Fig.~\ref{fig:dumping}.
This can be achieved using commonly available materials.
In this context, we utilize simple foam, typically used for packaging, and secure it to the cable
with two metallic plates, screws, and nuts.
We observe a significant decrease in the phase fluctuations,
as demonstrated in Fig.~\ref{fig:phase_vibrations} (blue curve).\\

@@

Electronics utilized to control the DMD chip experience thermal variations during operation.
The dynamics of this effect are dependent on the frame rate.
Specifically, the chips heats up more quickly when increasing the frame rate.
The increase of temperature can reach more that $15$ degrees Celsius
when running a sequence at maximum speed (20-30 kHz)~\cite{Rudolf2021thermal}.
Notably, this effect is less pronounced when the device is on but not running a sequence.
Temperature fluctuations can cause deformations on the chip's surface
and can modificy the phase response of the glass protective window.
This creates low order aberrations, which degrade the quality of the wavefront.
While this issue is comparatively less critical than static aberrations and mechanical instabilities
previously detailed,
it nonetheless has a substantial impact on the complex medium's response
when a DMD is employed to modulate the input wavefront.\\

Before initiating a wavefront shaping experiment,
it is important to characterize the influence of temperature
to assess the extent to which it impacts the results.
Although the exact effect on the wavefront distortion can be directly quantified~\cite{Rudolf2021thermal},
it is typically more convenient
to directly measure the effect on the studied system's response.
To do so, we use a setup to the one presented in the previous section
and depicted in Fig.~\ref{fig:MMF_ref}.
We can then estimate the field or intensity decorrelation over time.
We present here results with intesity correlation, as it does not require a reference arm.
The measured output pattern typically take the form of a seemingly random speckle pattern,
that is sensitive to minute changes of the input wavefront.
The correlation estimation is obtained by comparing the output intensity pattern
$I(\vec{r}, t)$ at a given time $t$
to the one at $t=0$.
We use the following expression for the correlation:

@@

The more efficient way to counteract this effect is to use a
close-loop system to stabilize the temperature of the DMD chip.
This can be achieved by using a thermoelectric cooler
as demonstrated in Ref.~\cite{Rudolf2021thermal}
and depicted in Fig.~\ref{fig:decorr_T}.

@@

\begin{tldr}
  \textbf{TL;DR:}\\
  DMDs take about an hour to thermaly stabilize when a sequence is running.
  It can conuntered by using a thermoelectric cooler
  and a temperature sensor.
  It can also be simply mitigated by letting a sequence running for few hours
  before starting the experiment.
\end{tldr}

>>>>

\documentclass[12pt]{iopart}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

@@

Digital Micromirror Devices (DMDs) have emerged as a technology bridging the gap
between these two types of systems;
they offer a large number of pixels (similar to LC-SLMs) and fast modulation speeds (typically up to several tens of kHz).
Their high speed capabilities made them attractive for real-time applications,
in particular for high-resolution imaging microscopy
requiring fast scanning or illumination shaping~\cite{Cha2000,Zhuang2020},
biolithography~\cite{yoon2018emerging},
and optical tweezers~\cite{Gauthier2016}.
However, DMDs are restricted to hardware binary amplitude modulation and are not optimized for coherent light applications.
Utilizing DMDs for coherent control of light in complex media is therefore non-trivial
and necessitates specific adaptations for efficient use.\\

To comprehend both the capabilities and limitations
of DMD technology for coherent wavefront shaping,
it is crucial to understand the device's operating principles
and its original design intentions.
Investigated and developed by Texas Instruments since the 1980s,
DMDs gained prominence in the 1990s for video projection applications since the 1990s
under the commercial name of Digital Light Processing (DLP)~\cite{Hornbeck1997,Dudley2003emerging}.
The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
DMDs operate by toggling the state of small mirrors between two distinct angles, denoted as $\pm \theta_\text{DMD}$.
The device is originally engineered for amplitude modulation in video projection applications.
In this configuration, one mirror angle directs light into the projection lens,
while the alternate angle results in the light path being blocked (see Fig.~\ref{fig:combined_pixel}).
Given that projectors utilize incoherent light and that the DMD plane is optically conjugated with the projection screen,
aberrations within the DMD plane are generally not problematic.
Similarly, phase fluctuations induced by temperature variations,
as well as minor vibrations from the cooling hardware, are inconsequential in this context.
The DMD is designed to produce binary on/off modulation,
which is then leveraged to generate grayscale images via pulse-width modulation.
Color modulation is accomplished through the use of a color wheel in conjunction with a bright white light source.\\

@@

The 2D array can be seen as 2 orthogonal 1D gratings in the $x$ and $y$ directions,
both having the same properties.
A more comprehensive depiction of the 2D system can be found in~\cite{Scholes2019structured}.
In order to utilize Eq.(\ref{eq:blazed_eq}),
one first needs to project
the different angles of the problem onto the incident planes
of the two 1D gratings.
This yields
$\alpha = \text{arctan}\left(\text{tan}(\alpha')/\sqrt{2}\right)$,
$\beta = \text{arctan}\left(\text{tan}(\beta')/\sqrt{2}\right)$,
and $\theta_B = \text{arctan}\left(\text{tan}(\theta_\text{DMD})/\sqrt{2}\right)$,
where $\theta_\text{DMD}$ represents the true rotation angle of the mirrors relative to the diagonal of the pixels,
and $\alpha'$ and $\beta'$ denote the incident and outgoing angles in the horizontal plane.
We can quantify how close we are to the ideal case,
i.e., when satisfying the blazed equation outlined by Eq.(\ref{eq:blazed_eq}),
by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:

@@

with $\lfloor . \rfloor$ representing the integer part,
and $\mod{2}$ the modulo 2 operation.
$\mu$ is maximal and equals $1$ when the blazing equation is satisfied,
i.e. when one order of diffraction contains most of the energy,
and minimal when we are in the worst-case scenario,
i.e. when four orders of diffraction have a significant and equal intensity.\\

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/mu_76.pdf}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width = \textwidth]{images/mu_108.pdf}
  \end{subfigure}
  \caption{
    \textbf{Blazed number and far field diffraction patterns, extreme cases.}
    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha'$  (top)
    for a pixel pitch of $d=7.6$\textmu m (left)
    and $d=10.8$\textmu m (right).
    Corresponding far field diffraction pattern (bottom)
    for two incident angles $\alpha' = 0^\circ$ and $\alpha' = 25^\circ$.
    The yellow cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
  }
  \label{fig:mu76}
\end{figure}

@@

\begin{tldr}
  \textbf{TL;DR:}\\
  DMDs act as blazed gratings.
  For a given operation wavelength,
  we need to find the right pixel pitch
  to have a good modulation quality
  and diffraction efficiency.
  It can be done by estimating the \textit{blazed number} $\mu$ introduced in eq.~(\ref{eq:blazed_number})
  or directly using our custom online tool~\cite{popoffDMDDiffractionTool}.
\end{tldr}

@@

While only capable of providing hardware binary amplitude modulation,
DMDs serve as a potent tool for wavefront shaping and sensing.
These applications  critically require characterization and correction of aberrations
caused by the non-flatness of the DMD surface.
For LC-SLMs, the manufacturer typically characterizes the surface's inhomogeneities within the plane of the modulator and
provides a spatial phase profile of the introduced aberrations based on the operational wavelength.
It is noteworthy that when utilized for intensity modulation on a plane conjugated to the DMD plane,
as it is done in digital projectors,
the system becomes insensitive to the aberrations caused by the DMD surface.
Consequently, these effects are commonly overlooked and rarely documented in the information provided by the manufacturers.
Nonetheless, they are non-negligible, with deviations from a flat surface
that typically spans several wavelengths~\cite{Brown2021multicolor}.\\

@@

In the literature, various methods have been proposed to characterize
the phase pattern of the DMD aberrations in the plane of the modulator.
Typically, this involves using a model for the aberrations,
tweaking the parameters to align with the measurements~\cite{Matthes2019Optical,Scholes2019structured, Brown2021multicolor}.
An alternative approach entails direct measurement of the distorted wavefront,
either employing a wavefront sensor such as a Shack-Hartmann~\cite{Lee2023compensation},
or via interferometry.
While such methods yield accurate results,
they necessitate adaptation to the particular
setup conditions,
and frequently require supplementary optical components,
meticulous alignment,
and custom software.
Furthermore, they also demand
a meticulously calibrated and stable setup.
In this section, we introduce a straightforward method to characterize the aberrations using a lens and a camera.
This technique can be employed for any system that offers phase modulation,
such as LC-SLMs or deformable mirrors. \\

@@

We assume the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}.
This implies that the modulator can be divided into $N$ sections,
which we designate as {\em macropixels},
where the phase can be controlled independently.
We use a lens and a camera in its Fourier plane,
illuminating the modulator with a collimated beam
that extends over the entire area of the modulator intended for use.
In scenarios where there are minimal or no aberrations,
the intensity pattern observed would mimic the PSF of the lens,
such as an Airy disk depicted in Fig.~\ref{fig:dmd_aberr_setup}.c.
However, in practice, we encounter a substantially distorted pattern,
like the one represented in Fig.~\ref{fig:dmd_aberr_setup}.d.
A more detailed depiction of the setup for aberration characterization,
in the context of a wavefront shaping application in complex media is presented in Fig.~\ref{fig:dmd_aberr_setup2}.a.
In the case where the medium is a multimode fiber,
one could leverage the reflection from the input surface
to directly visualize the intensity pattern of the input plane,
as shown in Fig.~\ref{fig:dmd_aberr_setup2}.b.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/aberr_setup.pdf}
  \caption{
    \textbf{Detailed aberration characterization setup.}
    (a) A laser beam is expanded and collimated using a telescope (lenses L and L').
    The DMD is represented in transmission for simplicity.
    Light reflected from the DMD is focused by a lens L$_\text{1}$
    onto the complex medium to study, namely a scattering medium (SC)
    or a multimode fiber (MMF).
    A beamsplitter is used to image the far field of the DMD onto a camera
    using a lens (L$_\text{2}$).
    The intensity pattern, up to a homotetic transformation,
    represent the input excitation on the medium
    which also corresponds to the PSF of the illumination setup.
    (b) In the case of a multimode fiber,
    one can take advantage of the reflection of the input facet
    to directly image the intensity pattern of the input plane.
  }
  \label{fig:dmd_aberr_setup2}
\end{figure}

@@

As an illustration, we conduct an optimization procedure using 11 Zernike polynomials.
We use a V-9501 Vialux DMD with a DLP9500 TI chip
of resolution 1920 by 1200 pixels and a pixel pitch of 10.8 \textmu m.
The optimization is performed on a disk of radius $340$ pixels,
The illumination is done using an expanded laser beam at 633 nm,
corresponding to the aperture of our optical setup.
We exclude the first three Zernike polynomials in the optimization process,
starting from the radial degree $2$.
Indeed, the initial one, known as the piston, does not influence the PSF quality.
The subsequent ones, the tip and the tilt, cause the PSF to shift.
Our procedure relies on optimizing the maximum of the PSF, wherever that is,
rendering us indifferent to these two parameters.
After optimization,
it is possible to generate the correction pattern using a selected number of Zernike polynomials
in order to investigate their impact on the PSF quality.
We see here that using about 10 Zernike polynomials is sufficient to obtain a Strehl ratio $>0.99$.
Fig.~\ref{fig:zernike} demonstrates the Strehl ratio and the intensity profiles of the PSF
for different counts of the utilized Zernike polynomials.
It is important to note that we do not use the full surface of the DMD.
Using a larger area may lead to stronger deformations of the PSF,
requiring a larger number of Zernike polynomials to be corrected accurately.
The experimental data, in addition to the Python code used to generate the figures, can be accessed in the dedicated repository~\cite{github}.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/Zernike_2.pdf}
  \caption{
    \textbf{Effect of Zernike polynomials on the optimization of the PSF.}
    (a) Strehl ratio of the point spread function as a function of the number of Zernike polynomials used
    for the compensation of the aberrations.
    (b) Intensity profile of the point spread function for different number of Zernike polynomials
    in linear and logarithmic scale.
    To generate this data, we take the final solution of the optimization experimental procedure,
    i.e. the optimal coefficients $a'_n$,
    and generate the corrected PSF with an increasing number of Zernike polynomials.
  }
  \label{fig:zernike}
\end{figure}

@@

Unlike the original purpose of the DMD, i.e. amplitude modulation for video projectors,
typical scientific applications require a high stability of the generated wavefront.
This is particularly true for applications in complex media,
such as strongly scattering media or multimode fiber,
where a small change in the phase front can lead to a large change in the output intensity profile.
While LC-SLMs design has been improved and adapted to scientific applications
over the last decades,
DMD are still relatively new tools for wavefront shaping and sensing
and are prone to instabilities that need to be addressed by the user.
In this section, we present the effect of mechanical and thermal instabilities
and how to limit their impact on the wavefront quality with simple solutions.\\

@@

In the supplementary materials~\cite{}, we present an animation illustrating the dynamic pattern.
In off-axis holography,
the local transverse displacement of the fringes
is directly proportional to the phase,
with a displacement equivalent to the period of the fringes corresponding to $2\pi$~\cite{Cuche2000spatial}.
This permits us to estimate the fluctuation in phase over time at a given position of the output plane.
As illustrated in Fig.~\ref{fig:phase_vibrations} (depicted by the red curve),
the phase varies rapidly over time,
a fluctuation attributed to the rapid mechanical vibrations transmitted by the board.\\

@@

A simple yet effective solution consists in dampening the vibrations
at the flat cable's level by clamping it with a soft material,
as depicted in Fig.~\ref{fig:dumping}.
This can be achieved using commonly available materials.
In this context, we utilize simple foam, typically used for packaging, and secure it to the cable
with two metallic plates, screws, and nuts.
We observe a significant decrease in the phase fluctuations,
as demonstrated in Fig.~\ref{fig:phase_vibrations} (blue curve).\\

@@

Electronics utilized to control the DMD chip experience thermal variations during operation.
The dynamics of this effect are dependent on the frame rate.
Specifically, the chips heats up more quickly when increasing the frame rate.
The increase of temperature can reach more than $15$ degrees Celsius
when running a sequence at maximum speed (20-30 kHz)~\cite{Rudolf2021thermal}.
Notably, this effect is less pronounced when the device is on but not running a sequence.
Temperature fluctuations can cause deformations on the chip's surface
and can modify the phase response of the glass protective window.
This creates low order aberrations, which degrade the quality of the wavefront.
While this issue is comparatively less critical than static aberrations and mechanical instabilities
previously detailed,
it nonetheless has a substantial impact on the complex medium's response
when a DMD is employed to modulate the input wavefront.\\

Before initiating a wavefront shaping experiment,
it is important to characterize the influence of temperature
to assess the extent to which it impacts the results.
Although the exact effect on the wavefront distortion can be directly quantified~\cite{Rudolf2021thermal},
it is typically more convenient
to directly measure the effect on the studied system's response.
To do so, we use a setup similar to the one presented in the previous section
and depicted in Fig.~\ref{fig:MMF_ref}.
We can then estimate the field or intensity decorrelation over time.
We present here results with intensity correlation, as it does not require a reference arm.
The measured output pattern typically take the form of a seemingly random speckle pattern,
that is sensitive to minute changes in the input wavefront.
The correlation estimation is obtained by comparing the output intensity pattern
$I(\vec{r}, t)$ at a given time $t$
to the one at $t=0$.
We use the following expression for the correlation:

@@

The more efficient way to counteract this effect is to use a
closed-loop system to stabilize the temperature of the DMD chip.
This can be achieved by using a thermoelectric cooler
as demonstrated in Ref.~\cite{Rudolf2021thermal}
and depicted in Fig.~\ref{fig:decorr_T}.

@@

\begin{tldr}
  \textbf{TL;DR:}\\
  DMDs take about an hour to thermaly stabilize when a sequence is running.
  It can countered by using a thermoelectric cooler
  and a temperature sensor.
  It can also be simply mitigated by letting a sequence run for few hours
  before starting the experiment.
\end{tldr}
